---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 320
)
```

# tidysmd

<!-- badges: start -->
[![R-CMD-check](https://github.com/malcolmbarrett/tidysmd/workflows/R-CMD-check/badge.svg)](https://github.com/malcolmbarrett/tidysmd/actions)
[![Codecov test coverage](https://codecov.io/gh/malcolmbarrett/tidysmd/branch/main/graph/badge.svg)](https://app.codecov.io/gh/malcolmbarrett/tidysmd?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/tidysmd)](https://CRAN.R-project.org/package=tidysmd)
<!-- badges: end -->

The goal of tidysmd is to easily create tidy data frames of SMDs. tidysmd wraps the smd package to easily calculate SMDs across many variables and using several weights in order to easily compare different adjustment strategies.

## Installation

You can install the most recent version of tidysmd from CRAN with:

``` r
install.packages("tidysmd")
```

Alternatively, you can install the development version of tidysmd from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("malcolmbarrett/tidysmd")
```

## Example

`tidy_smd()` supports both unweighted SMDs and weighted SMDs. 

```{r}
library(tidysmd)
tidy_smd(nhefs_weights, c(age, education, race), .group = qsmk)
```

`nhefs_weights` contains several types of propensity score weights for which we can calculate SMDs. Unweighted SMDs are also included by default.

```{r}
tidy_smd(
  nhefs_weights,
  c(age, race, education),
  .group = qsmk,
  .wts = c(w_ate, w_att, w_atm)
)
```

Having SMDs in a tidy format makes it easy to work with the estimates, for instance in creating Love plots. tidysmd includes `geom_love()` to make this a bit easier:

```{r}
library(ggplot2)
plot_df <- tidy_smd(
  nhefs_weights,
  race:active,
  .group = qsmk,
  .wts = starts_with("w_")
)

ggplot(plot_df) + geom_love()
```

