---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 320
)
```

# tidysmd

<!-- badges: start -->
[![R-CMD-check](https://github.com/malcolmbarrett/tidysmd/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/malcolmbarrett/tidysmd/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/malcolmbarrett/tidysmd/branch/main/graph/badge.svg)](https://app.codecov.io/gh/malcolmbarrett/tidysmd?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/tidysmd)](https://CRAN.R-project.org/package=tidysmd)
<!-- badges: end -->

The goal of tidysmd is to easily create tidy data frames of SMDs. tidysmd wraps the smd package to easily calculate SMDs across many variables and using several weights in order to easily compare different adjustment strategies.

## Installation

You can install the most recent version of tidysmd from CRAN with:

``` r
install.packages("tidysmd")
```

Alternatively, you can install the development version of tidysmd from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("malcolmbarrett/tidysmd")
```

## Example: Weighting

`tidy_smd()` supports both unweighted SMDs and weighted SMDs. 

```{r}
library(tidysmd)
tidy_smd(nhefs_weights, c(age, education, race), .group = qsmk)
```

`nhefs_weights` contains several types of propensity score weights for which we can calculate SMDs. Unweighted SMDs are also included by default.

```{r}
tidy_smd(
  nhefs_weights,
  c(age, race, education),
  .group = qsmk,
  .wts = c(w_ate, w_att, w_atm)
)
```

Having SMDs in a tidy format makes it easy to work with the estimates, for instance in creating Love plots. tidysmd includes `geom_love()` to make this a bit easier:

```{r}
library(ggplot2)
plot_df <- tidy_smd(
  nhefs_weights,
  race:active,
  .group = qsmk,
  .wts = starts_with("w_")
)

ggplot(
  plot_df,
  aes(
    x = abs(smd),
    y = variable,
    group = method,
    color = method,
    fill = method
  )
) +
  geom_love()
```

You can also use the quick-plotting function `love_plot()`, if you prefer:

```{r}
love_plot(plot_df) + 
  theme_minimal(14) + 
  ylab(NULL)
```

## Example: Matching

tidysmd also has support for working with matched datasets. Consider these two objects from the [MatchIt](https://github.com/kosukeimai/MatchIt) documentation:

```{r}
library(MatchIt)
# Default: 1:1 NN PS matching w/o replacement
m.out1 <- matchit(treat ~ age + educ + race + nodegree +
                   married + re74 + re75, data = lalonde)

# 1:1 NN Mahalanobis distance matching w/ replacement and
# exact matching on married and race
m.out2 <- matchit(treat ~ age + educ + race + nodegree +
                   married + re74 + re75, data = lalonde,
                   distance = "mahalanobis", replace = TRUE,
                  exact = ~ married + race)
```

One option is to just look at the matched dataset with tidysmd:

```{r}
matched_data <- get_matches(m.out1)

match_smd <- tidy_smd(
  matched_data,
  c(age, educ, race, nodegree, married, re74, re75), 
  .group = treat
)

love_plot(match_smd)
```

The downside here is that you can't compare multiple matching strategies to the observed dataset; the label on the plot is also wrong. tidysmd comes with a helper function, `bind_matches()`, that creates a dataset more appropriate for this task:

```{r}
matches <- bind_matches(lalonde, m.out1, m.out2)
head(matches)
```

`matches` includes an binary variable for each `matchit` object which indicates if the row was included in the match or not. Since downweighting to 0 is equivalent to filtering the datasets to the matches, we can more easily compare multiple matched datasets with `.wts`:

```{r}
many_matched_smds <- tidy_smd(
  matches,
  c(age, educ, race, nodegree, married, re74, re75), 
  .group = treat, 
  .wts = c(m.out1, m.out2)
) 

love_plot(many_matched_smds)
```





